\documentclass{aa} 

\usepackage{graphicx}
\usepackage[varg]{txfonts}
\usepackage{amsmath}

\pagestyle{plain}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\begin{document}  

\twocolumn[{
\begin{center}
  {\Large \bf FYS4150 Project 1: \\1-dimensional Poisson equation}\\[4ex]       
  {\large \bf Marie Foss$^{1}$, 
              Maria Hammerstr{{\o}}m$^{1}$}\\[4ex]
  \begin{minipage}[t]{15cm}
        $^1$ Institute of Theoretical Astrophysics, University of Oslo\\

  {\bf Abstract:}  We discovered \ldots 

  \vspace*{2ex}
  \end{minipage}
\end{center}
}] 


\section{Introduction}\label{sec:introduction}
In this project we will solve the one-dimensional Poissson equation with Dirichlet boundary conditions by rewriting it as a set of linear equations and solving these numerically in a program written in C++.The equation to be solved is:

\begin{equation}\label{eq:Poisson}
	-u''(x) = f(x) \hspace{0.5cm} x\in(0,1), \hspace{0.5cm} u(0) = u(1) = 0
\end{equation}

\noindent and we define the discretized approximation  to $u$ as $v_i$  with grid points $x_i=ih$   in the interval from $x_0=0$ to $x_{n+1}=1$. The step length or spacing is defined as $h=1/(n+1)$. We then have the boundary conditions $v_0 = v_{n+1} = 0$. We can approximate the second derivative of $u$ with:

\begin{equation}\label{eq:second_derivative}
   -\frac{v_{i+1}+v_{i-1}-2v_i}{h^2} = f_i  \hspace{0.5cm} \mathrm{for} \hspace{0.1cm} i=1,\dots, n,
\end{equation}

\noindent where $f_i=f(x_i)$.\\

\noindent Eq.~\ref{eq:second_derivative} can be written as a linear set of equations of the form: 

\begin{equation}
	{\bf A}{\bf v} = \tilde{{\bf b}}
\end{equation}

\noindent where $\tilde{b}_i=h^2f_i$ and \textbf{A} is an $n\times n$ matrix of the form:

\begin{equation}
    {\bf A} = \left(\begin{array}{cccccc}
                           2& -1& 0 &\dots   & \dots &0 \\
                           -1 & 2 & -1 &0 &\dots &\dots \\
                           0&-1 &2 & -1 & 0 & \dots \\
                           & \dots   & \dots &\dots   &\dots & \dots \\
                           0&\dots   &  &-1 &2& -1 \\
                           0&\dots    &  & 0  &-1 & 2 \\
                      \end{array} \right)
\end{equation}

\noindent (Show this!!) \\

\noindent We will assume that the source term is $f(x) = 100e^{-10x}$. Using the interval and boundary conditions as stated above, the above differential equation has a closed-form  solution given by $u(x) = 1-(1-e^{-10})x-e^{-10x}$. We will compare our numerical solution with this result. 





\section{Solving the problem}\label{sec:observations}
\subsection{Simple algorithm}
In our case we are dealing with a simple tridiagonal matrix. We can therefore rewrite our matrix \textbf{A} in terms of one-dimensional vectors \textit{a, b, c} of length 1:\textit{n}. The linear equation then reads:

\begin{equation}
    {\bf A} = \left(\begin{array}{cccccc}
                           b_1& c_1 & 0 &\dots   & \dots &\dots \\
                           a_2 & b_2 & c_2 &\dots &\dots &\dots \\
                           & a_3 & b_3 & c_3 & \dots & \dots \\
                           & \dots   & \dots &\dots   &\dots & \dots \\
                           &   &  &a_{n-2}  &b_{n-1}& c_{n-1} \\
                           &    &  &   &a_n & b_n \\
                      \end{array} \right)\left(\begin{array}{c}
                           v_1\\
                           v_2\\
                           \dots \\
                          \dots  \\
                          \dots \\
                           v_n\\
                      \end{array} \right)
  =\left(\begin{array}{c}
                           \tilde{b}_1\\
                           \tilde{b}_2\\
                           \dots \\
                           \dots \\
                          \dots \\
                           \tilde{b}_n\\
                      \end{array} \right).
\end{equation}

\noindent which can be written as:

\begin{equation}\label{eq:tridiagonal_system}
  a_iv_{i-1}+b_iv_i+c_iv_{i+1} = \tilde{b}_i
\end{equation}

\noindent for $i=1,2,\dots,n$. We want to find a simple algorithm to solve this set of equations. To do this, we can rewrite Eq.~\ref{eq:tridiagonal_system} by a method of elimination, which will lead to a system where there is only one unknown. 

The first thing to do is to realize that Eq.~\ref{eq:tridiagonal_system} gives a system of three equations:

\begin{align}
	b_1 v_1 + c_1 v_2 &= \tilde{b}_1, \hspace{0.5cm} i = 1  \\
	a_i v_{i-1} + b_i v_i + c_i v_{i+1} &= \tilde{b}_i, \hspace{0.5cm} i = 2, \dots, n - 1 \\
	a_n v_{n-1} + b_n v_n &= \tilde{b}_n, \hspace{0.5cm} i = n  
\end{align}


\noindent where the boundary conditions have been applied to simplify the equations. The idea now is to subtract one row with a scalar multiple of another to eliminate variables. First we want to eliminate $v_1$:\\

(something)\\

\noindent Then we can follow the same approac to eliminate $v_2$:\\

(something)\\

\noindent The algorithm for solving this equation can be stated as follows:\\

(algorithm)



\subsection{Using a library package}
In addition to solving the linear second-order differential equation Eq.~\ref{eq:Poisson} using the simple algorithm described above, we want to solve the equation using Gaussian elimination and LU decomposition, then compare the results.

More coming here.




\section{Analysis}    \label{sec:analysis}
\subsection{\dots}
\dots

\subsection{Number of floating point operations}
Some of the methods described above may be more computationally heavy for a computer to do than others. One way to evaluate this is to calculate the precise number of floating point operations ("FLOPS") needed to solve the equations.

For Eq.~\ref{eq:tridiagonal_system} we have that: FLOPS = $2 n^3$. For Gaussian elimination we have XXX and for LU decomposition XXX. 



\section{Conclusions} \label{sec:conclusions}
\dots




\end{document}

